Document Embedding for Scientific Articles:\newline A validation of word embeddings

### 2018-07-24 Paper/Parts/Appendix/appendix.tex
new paper start

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
\header{Research results}
In this chapter the results of our research are presented; the detailed discussion on the meaning and implications of these results is presented in Chapter 6: Discussion.
\subheader{Ranking}
The Figures~\ref{figure:titleRanks} \&~\ref{figure:abstractRanks} show the result of the categorization task as ranking results. The rank indicates the position of the correct journal in the sorted list of matched journals. Figure~\ref{figure:titleRanks} shows the ranking results for the different sets based on the title. Figure~\ref{figure:abstractRanks} displays the ranking results based on the abstract. Both graphs show both average and median ranks, based on the cosine-similarity between the article and journal embeddings or feature vectors.
\clearpage
\begin{figure}[hbt]
\includegraphics[width=6.5in]{Plots/Title_rankings}
\caption{Median and average title rankings}\label{figure:titleRanks}
\end{figure}
\begin{figure}[hbt]
\includegraphics[width=6.5in]{Plots/Abstract_rankings}
\caption{Median and average abstract rankings}\label{figure:abstractRanks}
\end{figure}
\clearpage
\subheader{Rank distribution}
Figures~\ref{figure:titleDistribution} and~\ref{figure:abstractDistribution} show the distributions of the ranks for each set. The Figures plot the summed amount of articles against the ranks on a logarithmic scale. Figure~\ref{figure:titleDistribution} shows the rank distribution for the titles, Figure~\ref{figure:abstractDistribution} shows this for the ranks based on the abstract. These graphs give a detailed view of the ranks presented in their respective Figures~\ref{figure:titleRanks} and~\ref{figure:abstractRanks}. The Y-axis of the two Figures shows the fraction of the total number of articles. Due to this, all sets will have a maximum of 1. When 1 is hit, the graph stagnates since the dataset has reached the maximum fraction. The X-axis plots the ranks of the articles. These ranks are the ranks of each individual article in our validation set, in contrast to Figure~\ref{figure:titleRanks} and~\ref{figure:abstractRanks} which show average and median ranks.
\begin{figure}[hbt]
\includegraphics[width=6.5in]{Plots/Title_rank_distribution}
\caption{Title rank distribution per set}\label{figure:titleDistribution}
\end{figure}
\begin{figure}[hbt]
\includegraphics[width=6.5in]{Plots/Abstract_rank_distribution}
\caption{Abstract rank distribution per set}\label{figure:abstractDistribution}
\end{figure}

\clearpage
\subheader{F1-Score}
Figures~\ref{figure:f1Title} and~\ref{figure:f1Abstract} show the precision, recall and F1 scores. These scores are calculated on journal level, and are averaged per set. Figure~\ref{figure:f1Title} shows the F1 score for the title and Figure~\ref{figure:f1Abstract} shows the scores for the abstract. These scores indicate the performance of the sets on absolute hits/top-1. We show the precision and recall scores since they vary between sets that have similar F1-scores (i.e. Figure~\ref{figure:f1Title}, tfidf\_5k\_5k and tfidf\_embedding
\begin{figure}[hbt]
\includegraphics[width=6.5in]{Plots/Title_avg_f1}
\caption{Precision, recall and F1 scores based on title}\label{figure:f1Title}
\end{figure}
\begin{figure}[htb]
\includegraphics[width=6.5in]{Plots/Abstract_avg_f1}
\caption{Precision, recall and F1 scores based on abstract}\label{figure:f1Abstract}
\end{figure}
\clearpage
\subheader{Memory usage}
Table~\ref{table:memoryUsage} shows the total memory usage of each set for the \textit{Validation set}, indicating their storage costs in gigabytes\footnote{1024 based}. It furthermore shows the absolute hit percentage of the title and the abstracts, i.e. the percentage of articles that have their source\footnote{Journal from which the article was taken} journal as the first result in the ranking. The table furthermore shows the median rank and the median abstract rank, as visualized in Figures \ref{figure:titleRanks} and~\ref{figure:abstractRanks}. Thus, this table gives an overview of the memory usage of the sets, combined with their performance on the ranking task.\\
\begin{table}[hbt]
\begin{center}
\begin{tabular}{|l|r|r|r|r|}
\hline
Set & Size in GB & Absolute hit percentage & Median title rank & Median abstract rank\\
&&\begin{tabular}{p{0.5in}|p{0.5in}}Title & Abstract\end{tabular}&&\\
\hline
tfidf 5k 5K & $9.82$ & \begin{tabular}{p{0.5in}|p{0.5in}}5.42\% & 10.18\%\end{tabular} & 50 & 27\\
\hline
tfidf 5K 10K & $11.47$ & \begin{tabular}{p{0.5in}|p{0.5in}}6.49\% & 11.08\%\end{tabular} & 38 & 15\\
\hline
\textbf{tfidf 10K 10K} &\textbf{11.61}& \begin{tabular}{p{0.5in}|p{0.5in}}\textbf{6.79\%} & \textbf{11.32\%}\end{tabular} & \textbf{35} & \textbf{14}\\
\hline
embedding & $3.13$ & \begin{tabular}{p{0.5in}|p{0.5in}}7.92\% & 9.24\%\end{tabular} & 27 & 23\\
\hline
5k embedding & $3.13$ & \begin{tabular}{p{0.5in}|p{0.5in}}6.34\% & 8.36\%\end{tabular} & 42 & 27\\
\hline
10k embedding & $3.13$ & \begin{tabular}{p{0.5in}|p{0.5in}}7.03\% & 8.76\%\end{tabular} & 34 & 25\\
\hline
\textbf{tfidf embedding} & \textbf{3.13} & \begin{tabular}{p{0.5in}|p{0.5in}}\textbf{7.89\%} & \textbf{9.33\%}\end{tabular} & \textbf{27} & \textbf{22}\\
\hline
1k 6k embedding & $3.06$ & \begin{tabular}{p{0.5in}|p{0.5in}}5.16\% & 7.86\%\end{tabular} & 64 &31 \\
\hline
\end{tabular}
\end{center}
\caption{Memory usage and performance for each set}\label{table:memoryUsage}
\end{table}
\subheader{Journal relatedness plot}
We created four seperate journal plots, Figures~\ref{figure:titlePlotNormal} and~\ref{figure:titlePlotGrouped} show the title based embeddings, and Figures~\ref{figure:abstractPlotNormal} and~\ref{figure:abstractPlotGrouped} show the abstract based embeddings.
Figures~\ref{figure:titlePlotNormal} and~\ref{figure:abstractPlotNormal} show the two dimensional plots of the journal embeddings. The journals are color-marked by publisher. Red is Wiley, lime is Elsevier and blue is Springer Nature. The grey points are other or not-specified publishers.
Figures~\ref{figure:titlePlotGrouped} and~\ref{figure:abstractPlotGrouped} show the journal embeddings grouped by a k-means algorithm, creating 8 groups. The bottom right of the plot shows the names of the journals closest to the center of the group. While most names state the topic of the research field, two do not. These are \textit{RSC Advances}: chemical sciences, and \textit{Symmetry}:  research on symmetry phenomena wherever they occur in mathematical or scientific studies. The k-means algorithm ran on the 300-dimensional vectors, while the plot shows the 2-dimensional vectors. In the title-based Figures (\ref{figure:titlePlotNormal} \&~\ref{figure:titlePlotGrouped}) the journals \textit{PNAS} and \textit{Cell} overlap each others labels on the bottom right side of the graph. In the abstract-based Figures (\ref{figure:abstractPlotNormal} \&~\ref{figure:abstractPlotGrouped}) the journals \textit{PNAS} and \textit{Nature} overlap each others label on the bottom right side of the graph.
The digital versions of these plots can be found at:~\url{https://goo.gl/nCGMcu} (title-based) and~\url{https://goo.gl/LwYP6k} (abstract-based). The usage of this digital version is advised, since this interactive plot enables zooming and adjusting plot settings, which can give a broader insight into the results than the presented figures. Although the presented figures give sufficient insight for the purpose of this thesis.
\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[height=6.5in]{Plots/Journal_Plots/Title_normal}
\end{center}
\caption{Journal plot of title embeddings, grouped by publisher}\label{figure:titlePlotNormal}
\end{figure}
\end{landscape}

>>>>

\documentclass[../../paper.tex]{subfiles}
\begin{document}
\section{Appendix}
\subsection{Journal Plots}
We created four seperate journal plots, Figures~\ref{figure:titlePlotNormal} and~\ref{figure:titlePlotGrouped} show the title based embeddings, and Figures~\ref{figure:abstractPlotNormal} and~\ref{figure:abstractPlotGrouped} show the abstract based embeddings.
Figures~\ref{figure:titlePlotNormal} and~\ref{figure:abstractPlotNormal} show the two dimensional plots of the journal embeddings. The journals are color-marked by publisher. Red is Wiley, lime is Elsevier and blue is Springer Nature. The grey points are other or not-specified publishers.
Figures~\ref{figure:titlePlotGrouped} and~\ref{figure:abstractPlotGrouped} show the journal embeddings grouped by a k-means algorithm, creating 8 groups. The bottom right of the plot shows the names of the journals closest to the center of the group. While most names state the topic of the research field, two do not. These are \textit{RSC Advances}: chemical sciences, and \textit{Symmetry}:  research on symmetry phenomena wherever they occur in mathematical or scientific studies. The k-means algorithm ran on the 300-dimensional vectors, while the plot shows the 2-dimensional vectors. In the title-based Figures (\ref{figure:titlePlotNormal} \&~\ref{figure:titlePlotGrouped}) the journals \textit{PNAS} and \textit{Cell} overlap each others labels on the bottom right side of the graph. In the abstract-based Figures (\ref{figure:abstractPlotNormal} \&~\ref{figure:abstractPlotGrouped}) the journals \textit{PNAS} and \textit{Nature} overlap each others label on the bottom right side of the graph.
The digital versions of these plots can be found at:~\url{https://goo.gl/nCGMcu} (title-based) and~\url{https://goo.gl/LwYP6k} (abstract-based). The usage of this digital version is advised, since this interactive plot enables zooming and adjusting plot settings, which can give a broader insight into the results than the presented figures. Although the presented figures give sufficient insight for the purpose of this thesis.
\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[height=6.5in]{Plots/Journal_Plots/Title_normal}
\end{center}
\caption{Journal plot of title embeddings, grouped by publisher}\label{figure:titlePlotNormal}
\end{figure}
\end{landscape}

### 2018-07-24 Paper/Parts/Conclusion/Conclusion.tex
new paper start

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
\header{Conclusion}
This research shows that the article embeddings, created with word embeddings, perform better than the reasonable TF-IDF alternatives on our categorization task, based on article titles. The TF-IDF alternatives give better results than the embeddings based on abstracts. The performance of the embeddings have been improved by weighing them with the TF-IDF values on the word level, although this improvement cannot be seen as significant on our dataset. This improved embedding set results in a median rank decrease of 8 on the titles and a median rank increase of 8 on the abstract, compared to the best performing TF-IDF alternative.  The embedding also results in a memory decrease of 73.04\% compared to the best performing TF-IDF alternative, making it more viable to keep it in memory. The visualization of the journal embedding shows that similar journals are grouped together, indicating a preservation of relatedness between the journal embeddings. We thus come to the following conclusions:
\begin{enumerate}
\item{Article based embeddings perform better than TF-IDF on titles, i.e. small texts which contain limited information}
\item{TF-IDF performs better than article based embeddings on abstracts, i.e. larger texts which contain more information}
\item{Embeddings give a significant decrease in memory usage compared to TF-IDF}
\item{Visualization of the journal embeddings show that the embeddings capture and preserve subject relatedness when they are combined to create embeddings for larger texts}

>>>>

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
\section{Conclusion}
This research shows that the article embeddings, created with word embeddings, perform better than the reasonable TF-IDF alternatives on our categorization task, based on article titles. The TF-IDF alternatives give better results than the embeddings based on abstracts. The performance of the embeddings have been improved by weighing them with the TF-IDF values on the word level, although this improvement cannot be seen as significant on our dataset. This improved embedding set results in a median rank decrease of 8 on the titles and a median rank increase of 8 on the abstract, compared to the best performing TF-IDF alternative.  The embedding also results in a memory decrease of 73.04\% compared to the best performing TF-IDF alternative, making it more viable to keep it in memory. The visualization of the journal embedding shows that similar journals are grouped together, indicating a preservation of relatedness between the journal embeddings. We thus come to the following conclusions:
\begin{enumerate}
\item{Article based embeddings perform better than TF-IDF on titles, i.e. small texts which contain limited information}
\item{TF-IDF performs better than article based embeddings on abstracts, i.e. larger texts which contain more information}
\item{Embeddings give a significant decrease in memory usage compared to TF-IDF}
\item{Visualization of the journal embeddings show that the embeddings capture and preserve subject relatedness when they are combined to create embeddings for larger texts}

### 2018-07-24 Paper/Parts/Data/Data.tex
paper setup
new paper start

\documentclass[../../Thesis.tex]{subfiles}

@@

\header{Research Data}
\subheader{Corpus}
The dataset we used for this research consists of articles published in 2017 which have been published in journals that have (in 2017) at least 150 publications. This selection leads to a dataset of $1.391.543$ articles from $3.759$ journals. Details about the corpus can be found in table~\ref{table:corpusSize}.

@@

\clearpage
\subsubheader{Text properties}
\citet{wiegand2018word} state that the Pareto distribution offers a good fit to the word occurrences in natural language. Their work shows that, due to the additional parameters in the Pareto-III\footnote{The pareto distribution is a distribution in which recognizable by a long "tail" of low values and a quick and short ascend to the top values. More information: \url{https://en.wikipedia.org/wiki/Pareto_distribution}.}, the tail of the data fits better with the model than the Zipf\footnote{Distribution similar to the Pareto distribution, see: \url{https://en.wikipedia.org/wiki/Zipf\%27s_law}} model. The fit to the Pareto distribution shows the relation between the word occurrences rank and the actual occurrences. This kind of word-occurrences distribution holds for many texts, including the writings of William Shakespeare, scientific texts and novels\cite{thurner2015understanding}.
The word occurrences in our corpus also follow the pattern of a Pareto distribution as described by~\citet{wiegand2018word}. Our word and token distribution is visualized in Figure~\ref{figure:wordTokenOccurrence}, which displays the occurrences of the first 500 tokens of the corpus.

@@

\includegraphics[width=6.5in]{Plots/word_occurrences}
\caption{Word and token occurrences. The word counts pattern resembles a Pareto distribution; the tokenization reduced the number of high-frequency words. The y-axis shows the number of occurances for each word, the x-axis shows the occurence rank of the word, meaning the position of the word in a list sorted on word occurance.}\label{figure:wordTokenOccurrence}

@@

\subheader{Datasets}
For this research we used a (pre-made) tokenized dataset, created from the earlier described corpus, which reduces the total amount of words by 34\% (see table~\ref{table:corpusSize}). From this tokenized set, we created the embeddings and the TF-IDF feature vectors.
\subsubheader{Tokenization}
The following steps have been applied to the words from the text to create a tokenized set:
\begin{enumerate}
\item{Removed punctuation}
\item{Removed all non-ASCII characters}
\item{Transformed all characters to lower-case}
\item{Removed stop-words, as provided by the NLTK\footnote{Natural Language ToolKit, \url{https://www.nltk.org/}} library}
\item{Removed numbers}
\item{Stemmed all words, using the stemmer provided by the NLTK library}
\end{enumerate}
These transformations reduced our dataset by 34\%, resulting in a tokenized set of $186.962.354$ tokens.
\clearpage
\subsubheader{Embedding}
For this research, we reused the word embeddings created by \citet{Truong2017Thesis}. These embeddings have a vector dimension of 300, which is an industry default. They have been trained on the entire Elsevier corpus (titles and abstracts respectively), not limited to the subset we used for this research. To create higher-level embeddings, we average all component embeddings. Thus, to create article embeddings (higher level), we take the average of all normalized word embeddings (component) for that article. Journal embeddings are created by taking the average of all normalized article embeddings. We use the average to combine multiple embeddings into one because we determine the meaning of a larger text as the "average meaning" of all components. Since this meaning is represented as an embedding, we can average the embeddings to create the average meaning. We have used multiple embedding optimizations for this research.
\subsubsubheader{Default embedding}
The default embedding is created from the pre-trained word embeddings; no modifications have been applied to this set. We refer to this set in the figures as \textbf{embedding}.
\subsubsubheader{TF-IDF weighted embedding}
The TF-IDF weighted embedding sets, referred to as TF-IDF embedding, are the default word embeddings weighted with a TF-IDF score per word.\\
The TF-IDF is calculated with a raw token count, and a smoothed inverted document frequency, calculated as follows:\\
\begin{equation}
IDF = \log_{10}(\dfrac{|A|}{|A_t|})
\end{equation}
Where $|A|$ is the total count of articles and $|A_t|$ is the count of articles containing term t.
The article embeddings are a normalized summation of each word vector multiplied by it TF-IDF value. Since we take a sum of all words, the Term Frequency is embedded as the raw count of each word. We will refer to this embedding in the figures as \textbf{tfidf\_embedding}.
\subsubsubheader{10K TF-IDF embedding}
The 10K embedding set is generated similarity to the TF-IDF embedding, this version only uses the 10.000 most common tokens, reducing the number of tokens it uses. This set was created to see if the limitation to 10.000 tokens reduces the amount of noise\footnote{Noise in this context are rare word}, increasing the performance. We will refer to this embedding in the figures as the \textbf{10k\_embedding}.
\subsubsubheader{5K TF-IDF embedding}
The 5K TF-IDF embedding is the TF-IDF embedding set, limited to the 5.000 most common words. This set was created to limit the number of tokens more aggressively, and with that, cancel out more noise. We will refer to this embedding in the figures as \textbf{5k\_embedding}.
\subsubsubheader{1K-6K TF-IDF embedding}
The 1K-6K TF-IDF embedding is the TF-IDF embedding limited to the top 6.000 most common words, without the top 1.000 most common words. The rationale for this cutting is that common words will occur in many articles, creating noise, by cutting off the top 1.000 and cutting off everything below 6.000 we tried to reduce the noise by filtering common words. This cut results in a set of 5.000 tokens, which allows us to compare it to the 5K TF-IDF set. We will refer to this embedding in the figures as \textbf{1k\_6k\_embedding}. We refer to the \textbf{10k\_embedding}, \textbf{5k\_embedding} and the \textbf{1k\_6k\_embedding} as the \textbf{limited TF-IDF embeddings}, because these embedding use TF-IDF weighing, and have been limited in their vocabulary due to the cut-off's.
\clearpage
\subsubheader{TF-IDF}
To create the TF-IDF feature vectors, we used the TF-IDF model and a hasher from PySparks MlLib library\footnote{\url{http://spark.apache.org/docs/2.0.0/api/python/pyspark.mllib.html}.}. The TF-IDF feature vectors are created by hashing the tokens with the hasher, which has a set hash bucket size. These hashed values are passed on to the TF-IDF model, resulting in a feature vector. The vector dimensions of the feature vector equal the number of hash buckets. To limit the computational and storage expenses and to reduce noise by rare words, we limit our vocabulary size.
We label the TF-IDF configurations as follows: $vocabulary size/hash bucket size$. Furthermore, we denote 1.000 as 1K, since we deal with chosen values which can be exactly noted given this notation. This means that the set with a 10.000 vocabulary size and a 10.000 hash bucket size will be denoted as "10K/10K". We will refer to the TF-IDF configurations in our figures as "tfidf\_vocabulary size\_hash bucket size".

@@

Figure~\ref{figure:tfidfPerformance} shows the performance of TF-IDF on title and abstract as median rank, and Figure~\ref{figure:tfidfMemory} shows the storage size in gigabyte, of the 1k/1K, 4K/4K, 7K/7K and 10K/10K TF-IDF configurations. This plot shows that, while the required storage size stagnates, the performance on title also quickly stagnates; the performance of the abstracts show similar behaviour. The stagnation of the storage size is likely due to the fact that the terms which are added to the larger vocabulary sets occur less often than the other words (the cut-off is based on word occurrence, descending). Given this information, we have chosen to use the 10K/10K, 10K/5K and 5K/5K configurations to compare our embedding results to because these configurations can be kept in memory (storage size) while it is not likely that larger sets will give much improvement. The TF-IDF features are created on article level. We average the set of article feature vectors to create a journal feature vector, just as we did with the embeddings. 
\header{Pipeline}
\subheader{Research environment}
The research has been done in a python-databricks environment, which used Spark SQL, a library that offers tools to work with big-data. \citet{armbrust2015spark} state that Spark SQL lets programmers leverage the benefits of relational processing and lets SQL users call complex analytic libraries in Spark. This allows for much tighter integration of relational and procedural processing. The paper further states that Spark SQL makes it significantly simpler and more efficient to write data pipelines that mix relational and procedural processing while offering substantial speedups over previous SQL-on-Spark engines.\\
\subheader{Data processing}
We processed the TF-IDF sets and the embedding sets \footnote{See Chapter 3.2 Datasets.} via the same pipeline, using their common vector properties. Using the same pipeline ensures comparable results. The pipeline is set-up as follows:
\begin{enumerate}
\item{Create training and validation set}
\item{Create journal embeddings}
\item{Categorize validation articles}
\item{Calculate performance metrics}
\end{enumerate}
\subheader{Create training and validation set}
We split our initial set  $80\%$ - $20\%$. We use the $80\%$ set as the training set for the journal representations, and the $20\%$ set as the validation set for the journal representations. This split is based on a random number given to each article, ensuring that all datasets (i.e. tfidf\_10k\_10k, embedding) have the same (random) training and validation set.
\subheader{Create journal embeddings}
From our training set we create the journal embeddings, which are created for most sets\footnote{See chapter 3.2 Datasets.} by averaging the article embeddings or feature vectors.
\clearpage
\subheader{Categorize validation articles}
To categorize the articles, we calculate the distance between the title of the article, and the title of each journal from the validation set. We also do this for the abstract of the article and the abstract of each journal. During this process we record of:

@@

\item{Title-based-rank of the actual journal}
\item{Abstract-based-rank of the actual journal}
\item{Best scored journal on the abstract similarity}
\item{Best scored journal on the title similarity}
\item{Abstract similarity between the actual journal and the article}
\item{Title similarity between the actual journal and the article}

@@

\subsubheader{Distance metrics}
To calculate the distance between two vectors, cosine similarity is commonly used. We validated the quality of cosine similarity as a distance metric by comparing it to other similarity metrics available in the SciPy library\footnote{\url{https://www.scipy.org/}}. We calculate the similarities based on the normalized embeddings, and compared the distance metrics based on the default embedding set. Table~\ref{table:distanceMetrics} shows the results of this validation.\\

@@

Dice & 1995 & 1929 & 1995 & 1929  \\

@@

Hamming & 1995 & 1929 & 1995 & 1929  \\
Jaccard & 1995 & 1929 & 1995 & 1929  \\
Kulsinski & 1995 & 1929 & 1995 & 1929  \\

@@

Matching & 1995 & 1929 & 1995 & 1929  \\
Rogerstanimoto & 1995 & 1929 & 1995 & 1929  \\
Russellrao & 1995 & 1929 & 1995 & 1929  \\

@@

Sokalmichener & 1995 & 1929 & 1995 & 1929  \\
Sokalsneath & 1995 & 1929 & 1995 & 1929  \\

@@

Yule & 1995 & 1929 & 1995 & 1929  \\

@@

\caption{Distance metric performance for word embeddings on the categorization of academic texts}\label{table:distanceMetrics}

@@

These results show high similarity between cosine-based metrics (Cosine \& Correlation) and euclidean based metrics (Euclidean, Seuclidean \& Sqeuclidean). This similarity is expected, since the cosine and euclidean distances should yield the same results on normalized sets\footnote{See: \url{https://en.wikipedia.org/wiki/Cosine_similarity\#Properties}.}. The results show that some enhancement on the euclidean algorithm (Seuclidean) result in slightly improved results, although not significant. Also the Cityblock metric yields results close to the Cosine metric, only having a slightly worse performance, which is also not significant. Because of this, we will use the cosine-similarity as the distance metric, which will make our results easier to compare with other work, since the cosine-similarity is a commonly used similarity measure.
\subheader{Performance measurement}
We use multiple metrics to validate the performance of the embedding sets and TF-IDF sets on the categorization task. These metrics are:
\begin{enumerate}
\item{F1-score}
\item{Median \& average rank}
\item{Rank distribution}
\end{enumerate}
\subsubheader{F1-score}

@@

\begin{jumpin}
$True Positive = $ Articles that are correctly matched to the current journal\\
$False Positive = $ Articles that are incorrectly matched to other journals\\
$False Negative = $ Articles that are incorrectly matched to the current journal\\
\end{jumpin}
With these definitions, we calculate the Recall, Precision \& F1 as follows:\\
\begin{jumpin}
$Recall = \dfrac{|True Positive|}{|True Positive| + |False Negative|}$\vspace{0.1in}\\
$Precision = \dfrac{|True Positive|}{|True Positive| + |False Positive|}$\vspace{0.1in}\\
$F1 = \dfrac{2 * Precision * Recall}{Precision + Recall|}$
\end{jumpin}
here we denote the amount of articles in a class as $|class|$.
\subsubheader{Median \& average rank}
A dataset rank is calculated using the ranks of all article in that dataset. The median rank of a dataset indicates the point where 50\% of all articles are to the left of a virtual line and 50\% are to the right of this line. The median rank therefore indicates the position of the "normal/typical" article. This median rank differs from the average, since the median is not influenced by a single value. If the median is 20, it does not matter if an article rank is 25 or 250, it does matter however if an article rank is 19 or 21. For the average rank this is not the case, the average rank is influenced by single articles. Therefore, it does not represents the "normal/typical" article but it represents the exact average rank of all articles in the validation set, which is changed when an article rank is changed from 25 to 250.\\
We use the median rank to indicate at which rank the typical article would be ranked, based on its title and abstract. We also calculate the average score for the title and abstract of all articles. This gives us four scores: \textbf{median title rank, median abstract rank, average title rank \& average abstract rank.}
\clearpage
\subsubheader{Rank distribution}
To further analyse the ranking results, we plot the article rank distribution to get an indication of the ranking-landscape.
\subsubheader{Two-dimensional plot}
To create a plot, we transformed the 300-dimensional journal vectors into 2-dimensional vectors using TSNE based on PCA\footnote{Principal component analysis}. These 2-dimensional vectors, representing the x \& y coordinate, can then be drawn in a plot. To visualize the preservation of journal-relatedness while converting the 300 dimensions to 2 dimensions, we create groupings using k-means. These groupings are created on the 300-dimensional vectors and are visualized in the plot using colors. We use the k-means groupings due to the lack of subject-based grouping values in our dataset, for this research we used 2, 5, 8, 10, 20, 25, 30, 35 and 40 as the amount of groups. We included only the plots of 8 groups, because we find this plot the most informative. All group amounts can be seen in the digital version of the plots.

>>>>

\documentclass[../../paper.tex]{subfiles}

@@

\section{Research Data}
\subsection{Corpus}
The dataset we used for this research consists of $1.391.543$ articles from $3.759$ journals, all articles have been published in journals in 2017, that have (in 2017) at least 150 publications. Details about the corpus can be found in table~\ref{table:corpusSize}. The distribution of the text in our dataset follows a Pareto-like distribution, which is a common property of texts (\citet{wiegand2018word}). Our word and token distribution is visualized in Figure~\ref{figure:wordTokenOccurrence}, which displays the occurrences of the first 500 tokens and words of the corpus, ordered on occurrences.

@@

@@

\includegraphics[width=4.5in]{Plots/word_occurrences}
\caption{Word and token occurrences. The y-axis shows the number of occurances for each word/token, the x-axis shows the occurence rank of the word/token; the position of the word/token in a list sorted on occurance.}\label{figure:wordTokenOccurrence}

@@

\subsection{Datasets}
For this research we used a tokenized dataset, created from the earlier described corpus (see table~\ref{table:corpusSize}). From this tokenized set, we created the embeddings and the TF-IDF feature vectors. The following steps have been applied to convert the words from the text to tokens: removing punctuation, non-ascii characters, number \& stop-words, transforming all characters to lower-case and by stemming all remaining tokens using the NLTK stemmer\footnote{Natural Language ToolKit library, \url{https://www.nltk.org/}.}.
These steps reduced our dataset by 34\% (words to tokens), resulting in a set of $186.962.354$ tokens.
\subsubsection{Embedding}
For this research, we reused the word embeddings created by \citet{Truong2017Thesis}. These embeddings have a vector dimension of 300, and have been trained on the entire Elsevier corpus (titles and abstracts respectively), not limited to the subset we used for this research. To create higher-level embeddings, we average all component embeddings. Thus, to create article embeddings (higher level), we take the average of all normalized word embeddings (component) for that article. Journal embeddings are created by taking the average of all normalized article embeddings. We use the average to combine multiple embeddings into one because we determine the meaning of a larger text for this research as the "average meaning" of all components. We have used multiple embedding variations for this research. \textbf{Default embedding}; the embeddings  obtained from the word embeddings are referred to as the default embedding, \textit{embedding} in figures. A variation on this embedding is the \textbf{TF-IDF weighted embedding}. The word embeddings in this set have been weighted with the TF-IDF value for each word, we refer to this set as TF-IDF embedding, \}.{tfidf\_embedding} in figures. The TF-IDF score is calculated using the raw token count and the smoothed inverted document frequency. The \textbf{10K TF-IDF embedding} uses the TF-IDF embedding as basis, The 10K TF-IDF embedding set has been weighted with the same TF-IDF, but is limited to the top 10.000 most occurring tokens. This set is referred to as 10K TF-IDF embedding, \textit{10K\_embedding} in the figures. Another variation on the TF-IDF set is the \textbf{5K TF-IDF embedding}. The 5K TF-IDF embedding has also been weighted with the TF-IDF values, but has been limited to the top 5.000 most occurring tokens. This set is referred to as 5K TF-IDF embedding, \textit{5K\_embedding} in the figures. The \textbf{1K-6K TF-IDF embedding} set weights the word embeddings with the TF-IDF values and is limited to the tokens between the top 1.000 and top 6.000 based on their occurrence. Creating a set of 5.000 tokens, without using the top 1.000 tokens. We refer to this set as \textit{1k\_6k\_embedding} in the figures.
\subsubsection{TF-IDF}
The TF-IDF feature vectors are created using the TF-IDF model and a hasher from the PySpark MlLib\footnote{\url{http://spark.apache.org/docs/2.0.0/api/python/pyspark.mllib.html}.}.The TF-IDF feature vectors are created by hashing the tokens with the hasher, which has a set hash bucket size. These hashed values are passed on to the TF-IDF model, resulting in a feature vector with a dimension equal to the amount of hash buckets. To limit the computational and storage expenses and to reduce noise by rare words, we limit our vocabulary size. We label the TF-IDF configurations as follows: $vocabulary size/hash bucket size$. Furthermore, we denote 1.000 as 1K, since we deal with chosen values which can be exactly noted given this notation. This means that the set with a 10.000 vocabulary size and a 10.000 hash bucket size will be denoted as "10K/10K". We will refer to the TF-IDF configurations in our figures as "tfidf\_vocabulary size\_hash bucket size". Figure~\ref{figure:tfidfPerformance} shows the performance of TF-IDF on title and abstract as median rank, and Figure~\ref{figure:tfidfMemory} shows the storage size in gigabyte, of the 1k/1K, 4K/4K, 7K/7K and 10K/10K TF-IDF configurations. These figures show that, while the required storage size stagnates, the performance on title also quickly stagnates; the performance of the abstracts show similar behaviour. The stagnation of the storage size is likely due to the fact that the terms which are added to the larger vocabulary sets occur less often than the other words (the cut-off is based on word occurrence, descending). Given this information, we have chosen to use the 10K/10K, 10K/5K and 5K/5K configurations to compare our embedding results to because these configurations can be kept in memory (storage size) while it is not likely that larger sets will give much improvement. The TF-IDF feature vectors are created on article level and we average the set of article feature vectors to create a journal feature vector, just as we did with the embeddings.
\section{Pipeline}
\textbf{Data processing;} The research has been done in a python-databricks environment, which used Spark SQL, a library that offers tools to work with big-data. We processed the TF-IDF sets and the embedding sets \footnote{See Chapter 3.2 Datasets.} via the same pipeline, using their common vector properties. The pipeline contains the following steps: creating a training and validation set, creating the journal embeddings from the training set, categorizing the validation set articles using the journal embeddings and calculating the performance metrics.

@@

@@

\item{Create training and validation set;
We split our initial set in a training set ($80\%$) and a validation set ($20\%$). This split of the data is based on a random number given to each article, ensuring that all datasets (i.e. tfidf\_10k\_10k, embedding) have the same (random) training and validation set.}
\item{Create journal embeddings;
From our training set we create the journal embeddings, which are created for most sets\footnote{See chapter 3.2 Datasets.} by averaging the article embeddings or feature vectors.}
\item{Categorize validation articles;
To categorize the articles, we calculate the distance between the title of the article (validation set), and the title of each journal (training set). We also do this for the abstract of the article and the abstract of each journal. During this process we record of the title-based-rank and abstract-based-rank of the source journal, best scored journal on abstract and title and the similarity between the source journal and article for both abstract and title.}
\item{Distance metrics;
To calculate the distance between two vectors, cosine similarity is commonly used. We validated the quality of cosine similarity as a distance metric by comparing it to other similarity metrics available in the SciPy library\footnote{\url{https://www.scipy.org/}}. We calculate the similarities based on the normalized embeddings, and compared the distance metrics based on the default embedding set. Table~\ref{table:distanceMetrics} shows the results of this validation.}

@@

@@

@@

@@

@@

@@

@@

\caption{Distance metric performance for word embeddings on the categorization of academic texts, only showing relevant results.}\label{table:distanceMetrics}

@@

The results in Table~\ref{table:distanceMetrics} show high similarity between cosine-based metrics (Cosine \& Correlation) and euclidean based metrics (Euclidean, Seuclidean \& Sqeuclidean). This similarity is expected, since the cosine and euclidean distances should yield the same results on normalized sets\footnote{See: \url{https://en.wikipedia.org/wiki/Cosine_similarity\#Properties}.}. The results show that some enhancement on the euclidean algorithm (Seuclidean) result in slightly improved results, although not significant. Also the Cityblock metric yields results close to the Cosine metric, only having a slightly worse performance, which is also not significant. Because of this, we will use the cosine-similarity as the distance metric, which will make our results easier to compare with other work, since the cosine-similarity is a commonly used similarity measure.
\subsection{Performance measurement}
We use multiple metrics to validate the performance of the embedding sets and TF-IDF sets on the categorization task. These metrics are: the F1-score, the Median \& average rank and the Rank distribution.
\item{F1-score;

@@

$True Positive = $ Articles correctly matched to the current journal\\
$False Positive = $ Articles incorrectly matched to other journals\\
$False Negative = $ Articles incorrectly matched to the current journal\\
With these definitions, we calculate the Recall, Precision \& F1 using the standard formulas.}
\item{Median \& average rank;
We use both average and median ranks as rank-indication for a journal. The median rank indicates at which point 50\% of all values is larger and 50\% of all values is smaller, indicating the "typical article". While the average indicates the exact average of all articles in the set. We used both measures for our research.}
\item{Rank distribution;
To further analyse the ranking results, we plot the article rank distribution to get an indication of the ranking-landscape.}
